ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-09 pyked.tex
extended shock tube example, with figure

The \ck{} file used in this example can be found in the \texttt{tests} directory of the \pk{}
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in
Section~\ref{sec:rcm-modeling}.

This \ck{} file specifies multiple data points with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox|common-properties| section is specified, followed by the
\yabox|datapoints| list:
%
\begin{yamlbox}
common-properties:
  composition: &comp
    kind: mole fraction
    species:
      - species-name: t-butanol
        InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
        amount:
          - 0.003333333
      - species-name: O2
        InChI:  1S/O2/c1-2
        amount:
          - 0.04
      - species-name: Ar
        InChI:  1S/Ar
        amount:
          - 0.956666667
  ignition-type: &ign
    target: OH*
    type: 1/2 max
datapoints:
  - temperature:
      - 1459 kelvin
    ignition-delay:
      - 347 us
    pressure:
      - 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1389 kelvin
    ignition-delay:
      - 756 us
    pressure:
      - 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
    ...
\end{yamlbox}
%
As before, we have truncated the datapoints list for space.

In this example, we will run constant-volume simulations at each
pressure and temperature condition in the \yabox|datapoints| list. Once again,
the \ck{} file specifies all the information required for the simulations except
for the chemical kinetic model, and Cantera can be used to simulate autoignition. The setup steps
are the same as in the previous example, with one exception. In this example, we also import
Python's built-in multiprocessing library so that we can run the simulations on multiple cores. We
import the \pybox|Pool| class, which offers facilities to manage a jobs on a pool of processors:
%
\begin{pythonbox}
import cantera as ct
from multiprocessing import Pool
from pyked import ChemKED
\end{pythonbox}

Then, we define a function that will be mapped onto each job. This function takes the initial
temperature, pressure, and mole fractions as input and returns the simulated ignition delay time,
defined as the time the temperature increases by \SI{400}{\K} over the initial temperature, a
simplified definition for this example. In general, the user could process the mole fraction of
\ce{OH^*} (provided that the kinetic model includes accurate chemistry for \ce{OH^*}) to match the
definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
be supplied by the user:
%
\begin{pythonbox}
def run_simulation(T, P, X):
    gas = ct.Solution('tbuoh-mechanism.cti')
    gas.TPX = T, P, X
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    while reac.T < T + 400:
        netw.step()

@@

Finally, we create the processor \pybox|Pool| (with 4 processes) and send the jobs out to run:
%
\begin{pythonbox}
with Pool(processes=4) as pool:
    ignition_delays = pool.starmap(run_simulation, initial_conditions)

for (T, P, X), tau in zip(initial_conditions, ignition_delays):
    print(f'The ignition delay for T_initial={T} K, P_initial={P} Pa is: {tau} seconds')
\end{pythonbox}
%
The simulated ignition delay results are returned in the \pybox|ignition_delays| list. The results
are printed to the screen using a Python formatted string (f-string).
%====================================================================
\section{Conclusions and future work}

>>>>

The \ck{} file used in this example can be found in the \texttt{tests} directory of the \pk{}
repository.\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in
Section~\ref{sec:rcm-modeling}.
This \ck{} file specifies multiple data points with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox|common-properties| section is specified, followed by the
\yabox|datapoints| list (as before, we have truncated the \yabox|datapoints| list for space):
%
\begin{yamlbox}
common-properties:
  composition: &comp
    kind: mole fraction
    species:
      - species-name: t-butanol
        InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
        amount:
          - 0.003333333
      - species-name: O2
        InChI:  1S/O2/c1-2
        amount:
          - 0.04
      - species-name: Ar
        InChI:  1S/Ar
        amount:
          - 0.956666667
  ignition-type: &ign
    target: OH*
    type: 1/2 max
datapoints:
  - temperature:
      - 1459 kelvin
    ignition-delay:
      - 347 us
    pressure:
      - 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature:
      - 1389 kelvin
    ignition-delay:
      - 756 us
    pressure:
      - 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
    ...
\end{yamlbox}

In this example, we will run constant-volume simulations at each
pressure and temperature condition in the \yabox|datapoints| list. Once again,
the \ck{} file specifies all information required for the simulations except
for the chemical kinetic model, and Cantera can be used to simulate autoignition. The setup steps
match those from the previous example, with one exception: in this example, we also import
Python's built-in multiprocessing library so that we can run the simulations on multiple cores.
We import the \pybox|Pool| class, which offers tools to manage a set of jobs on a
pool of processors:
%
\begin{pythonbox}
import cantera as ct
from multiprocessing import Pool
from pyked import ChemKED
\end{pythonbox}

Then, we define a function that will be mapped onto each job. This function takes the initial
temperature, pressure, and mole fractions as input and returns the simulated ignition delay time,
defined as the time the temperature increases by \SI{400}{\K} over the initial temperature, a
simplified definition for this example. In general, the user could process the mole fraction of
\ce{OH^*} (provided that the kinetic model includes accurate chemistry for \ce{OH^*}) to match the
definition of ignition delay in the experiments. We use the chemical kinetic model
for butanol isomers from Sarathy et al.~\cite{Sarathy:2012fj}, represented as
\verb|LLNL_sarathy_butanol.cti|:
%
\begin{pythonbox}
def run_simulation(T, P, X):
    gas = ct.Solution('LLNL_sarathy_butanol.cti')
    gas.TPX = T, P, X
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    while reac.T < T + 400:
        netw.step()

@@

Finally, we create the processor \pybox|Pool| (with four processes) and send the jobs out to run:
%
\begin{pythonbox}
with Pool(processes=4) as pool:
    ignition_delays = pool.starmap(run_simulation, initial_conditions)

for (T, P, X), tau in zip(initial_conditions, ignition_delays):
    print(f'The ignition delay for T_initial={T} K, P_initial={P} Pa is: {tau} seconds')
\end{pythonbox}
%
The simulated ignition delay results are returned in the \pybox|ignition_delays| list. The results
are printed to the screen using a Python formatted string (f-string). We can also visually
compare the results using Matplotlib~\cite{matplotlib}, as
Figure~\ref{fig:ignition-delays} shows:
%
\begin{pythonbox}
import matplotlib.pyplot as plt

inv_temps = [1000/i[0] for i in initial_conditions]
exp_ignition_delays = [dp.ignition_delay.to('ms').magnitude for dp in ck.datapoints]
sim_ignition_delays = np.array(ignition_delays)*1.0E3

plt.figure()
plt.scatter(inv_temps, exp_ignition_delays, label='Experimental ignition delays')
plt.scatter(inv_temps, sim_ignition_delays, label='Simulated ignition delays', marker='s')
plt.legend(loc='best')
plt.yscale('log')
plt.ylabel('Ignition delay [ms]')
plt.xlabel('1000/T [1/K]')
\end{pythonbox}

\begin{figure}[htb]
    \centering
    \includegraphics[width=0.7\textwidth]{ignition-delays.pdf}
    \caption{Comparison of experimental and simulated ignition delays of
    \textit{tert}-butanol from shock tube example.}
    \label{fig:ignition-delays}
\end{figure}

%====================================================================
\section{Conclusions and future work}
